---
title: "Geography 176C"
subtitle: "NWM catchments and flow data"
author: "Angus Watters"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ucsb.css
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 2; font-size: 80%}
```
```{r, include = F}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)
library(flipbookr)
library(gapminder)
library(tidyverse)

climate_graph <- function(df, lab) {
  df$date <- paste0(df$date, "-01")
  df$date <- as.Date(df$date)
  rownames(df) <- df$date
  df <- select(df, 5)

  dygraph(data = df,
            ylab = as.character(lab)) %>% 
      dyHighlight(highlightCircleSize = 4,
                  highlightSeriesBackgroundAlpha = .4) %>%
      dyOptions(colors = c("navy"),
                fillGraph = TRUE)
}
```

# *Lightning talk!*

---

# **The National Water Model (NWM)**

- The National Water Model (NWM) is a hydrologic modelling framework that simulates observed and forecast streamflow over the entire continental United States (CONUS). 

- The NWM simulates the water cycle with mathematical representations of the different processes and how they fit together.

--

# **Motivation**

The NWM obtains data from over 8000 USGS stream gauges in the Nation and uses this data to expand forecasts to over 2.7 million stream locations nationwide.
This also means that there is a huge amount of historical streamflow data avalialbe at all these locations.

---

# **Problem**

There is not an easy way to view a particular catchment and see the historical streamflow records for this catchment. The data is all there but currently there isn't a quick and easy way to view both the catchment geometry and timeseries data on demand.  

--

# **Solution**
- Create a R Shiny dashboard that features an interactive map and allows the user to explore all watershed catchments in the US and to get timeseries streamflow data on demand by clicking on a catchment.

- This project will be created in R using the following packages: 

--

**R packages**  
`tidyr`
`dplyer`
`sf`
`leaflet`
`dataRetrieval`
`nhdplusTools`
`nwmTools`
`shiny`
---

# Get COMID from a lat/long point

`r flipbookr::chunk_reveal("nldi", title = "Find COMID from a lat/lng point")`
```{r nldi, include=FALSE}
library(dataRetrieval)

nldi <- findNLDI(location = c(-95.07, 30.36))
nldi 

```

---
# Get catchment polygons from a NHD COMID
`r flipbookr::chunk_reveal("nhd", title = "# Question 1: Counties with most cases")`
```{r nhd, include = FALSE}
library(nhdplusTools)
library(ggplot2)
nhd <- get_nhdplus(comid = nldi$comid, realization = "catchment")
nhd

ggplot() +
  geom_sf(data = nhd)
```
---

# Get timeseries data using COMID/point
`r flipbookr::chunk_reveal("nwm", title = "# Question 2: States with most cases")`

```{r nwm, include = FALSE}
library(climateR)
library(sf)
library(dygraphs)

pt <- st_centroid(nldi)
precip <- getTerraClim(AOI = pt, param = "prcp", startDate = "2010-01-01", endDate = "2014-01-01")
climate_graph(precip, "Volume (mm)")
```
---

# The hidden need for ungrouping

- `ungroup()` should always be  applied after the grouping calcuations.. 
- If you forget to `ungroup()` data, future data management will likely produce errors.
-  Even if you do not plan on performing additional calculations, it’s a good habit to keep. 
- `ungroup()` is especially important when creating objects!!

--

- Think about the dimision of your data structure!

--

.pull-left[
```{r}
gapminder %>% 
  filter(year == 2007) %>% 
  group_by(continent) %>% 
  mutate(mp = mean(pop)) %>%   
  mutate(mle = mean(lifeExp)) %>% 
  ungroup()  %>%  #<<
  select(country, year, mp, mle) 
```
]
